---
title: "R Notebook Example"
author: "Simon Schwab & Eva Furrer"
date: "30 Jan 2019"
output: html_notebook
---

# 1. Install required libraries (only once)
```{r}
# install.packages("carData")
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
# install.packages("ggplot2")
```

# 2. Load required libraries
```{r}
library(biostatUZH)
library(ggplot2)
```

# 3. Load data
```{r}
data <- carData::TitanicSurvival
```

# 4. Figure
Example figure with distribution of age per class.
```{r, fig.height=2, fig.width=2, warning=FALSE}
ggplot(data, aes(x=passengerClass, y=age, color=passengerClass)) + 
  geom_boxplot(width=0.8) +
  geom_point(shape=1, color="gray70", size=1, position=position_jitter(width=0.3, height=0.01)) +
  ylab("Age") + xlab("Passenger Class") + theme_minimal() + theme(legend.position="none") 
```


# Model binary outcome survived yes/no
```{r}
data$survived <- relevel(data$survived, ref="yes") # relevel: baseline is survived: yes

fit <- glm(survived ~ sex + passengerClass + age,
          data=data, family=binomial())
```

# Create table for publication
```{r}
labels <- c("female", "1st") # reference levels of the two categorial variables
tableOR(fit, latex=FALSE, short=T, refLevels=labels, rmStat=TRUE,
        caption="Changes in odds for risk of death in the Titanic tragedy.")
```

Male passengers had a 12-fold increase in the odds to die compared to female. Passengers from 3rd class have a almost 10-times increase in odds to die, and passengers from 2nd class have a 3.6-times increase in odds compatred to 1st class.