---
title: "R Notebook Example"
author: "Simon Schwab & Eva Furrer"
date: "14 Jan 2019"
output: html_notebook
---

# Install required libraries (only once)
```{r}
# install.packages("biostatUZH")
```

# Load required libraries
```{r}
library(biostatUZH)
library(ggplot2)
```

# Load & prepare data
```{r}
dat = carData::TitanicSurvival
dat$survived = relevel(dat$survived, ref = "yes") # relevel: baseline is survived: yes
```

# Figure with some descriptives
```{r}
ggplot(dat, aes(x=passengerClass, y=age, color=passengerClass)) +   geom_boxplot(width=0.8) +
  geom_point(shape=1, color="gray70", size=1, position = position_jitter(width = 0.3, height = 0.01))
```


# Model binary outcome survived yes/no
```{r}
fit = glm(survived ~ sex + age + passengerClass,
          data = dat, family = binomial())
summary(fit)
```

# Create table for publication
```{r}
labels = c("female", "1st") # reference levels of the two categorial variables
tableOR(fit, latex = FALSE, short = TRUE, refLevels = labels, caption = "Changes in odds for risk of death in the Titanic tragedy.")
```

Male passengers have a 12-fold increase in odds to die compared to female. Passengers from 3rd class have a almost 10-times increase in odds to die, and passengers from 2nd class have a 3.6-times increase in odds compatred to 1st class.